---
title: "Working with Vectors"
author: "Hailey Veirs"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## Creating Vectors
Vectors must have the same class, such as "chr" (character), "num" (numeric), "float" (floating), "int" (integers)
Can check class using "class()" function

```{r}
marmots <- c("blue", "green", 4, "yellow")

#Check the class of our vector
class(marmots)
```

```{r}
pika <- c(12.4, 6.8, 8.6, 7.4, 5.3)

class(pika)

# Do scalar multiplication on the pika vector
pika_scalar <- 5.2 * pika
```

Vector operations

Dot product is done using %*% between the vectors in the order to want the dot product to be.

```{r}
bananas <- c(1, 2, 3)
apples <- c(6, 7, 8)

bananas + apples
bananas - apples

# Dot product of vectors is "%*%"
bananas %*% apples
```

## Making Matrices 

To make a matrix, you can make a sequence of values and then tell R to make a matrix of m rows and n columns

```{r}
my_values <- seq(from = 1, to = 10) # can add "length" to make 50 diff numbers evenly spaced

my_matrix <- matrix(data = my_values, nrow = 2, ncol = 5, byrow = TRUE)

# Matrices operate similarly to vectors
5 * my_matrix # scalar multiplication
class(my_matrix)

# "glimpse" shows info on data frames and matricies
glimpse(penguins)

```

## Tasks 

### Section 1
